### 2014-05-12 writeup/paper.tex
Some paper edits, added some citations

% TEMPLATE for Usenix papers, specifically to meet requirements of
%  USENIX '05
% originally a template for producing IEEE-format articles using LaTeX.
%   written by Matthew Ward, CS Department, Worcester Polytechnic Institute.
% adapted by David Beazley for his excellent SWIG paper in Proceedings,
%   Tcl 96
% turned into a smartass generic template by De Clarke, with thanks to
%   both the above pioneers
% use at your own risk.  Complaints to /dev/null.
% make it two column with no page numbering, default is 10 point

% Munged by Fred Douglis <douglis@research.att.com> 10/97 to separate
% the .sty file from the LaTeX source template, so that people can
% more easily include the .sty file into an existing document.  Also
% changed to more closely follow the style guidelines as represented
% by the Word sample file.

% Note that since 2010, USENIX does not require endnotes. If you want
% foot of page notes, don't include the endnotes package in the 
% usepackage command, below.

% This version uses the latex2e styles, not the very ancient 2.09 stuff.
\documentclass[letterpaper,10pt]{article}
\usepackage{epsfig,graphicx,usenix,fullpage,float,hyperref}
\usepackage{mathtools,amssymb,setspace, pbox}
\usepackage[table,xcdraw]{xcolor}
\usepackage[center, labelfont=bf]{caption}

\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
\def\bE{\mathbb{E}}

%\usepackage{endnotes}
\begin{document}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf 6.824 Final Project}
%for single author (just remove % characters)
\author{
{\rm Colleen Josephson}\\
cjoseph@mit.edu
\and
{\rm Joseph DelPreto}\\
delpreto@mit.edu
\and
{\rm Pranjal Vachaspati}\\
pranjal@mit.edu
\and
{\rm Steven Valdez}\\
dvorak42@mit.edu
% copy the following lines to add more authors
% \and
% {\rm Name}\\
%Name Institution
} % end author

\date{May 11, 2014}

\maketitle

% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
%\thispagestyle{empty}

\section{Introduction}
The presented project implements a persistent, fault-tolerant, high-performance key/value store. In order to achieve these goals, the system implements at-most-once semantics, relies upon Paxos for reaching consensus, shards the database across multiple replica groups, persistently stores values to disk, and implements protocols for automatically recovering and updating disk contents upon server reboot.  In order to improve performance, the system features two paxos optimizations: a protocol similar to multi-paxos to avoid the dueling leaders problem, and the ability for leaders to send out prepare messages in advance to reduce the required number of RPCs for paxos agreement.

Extensive testing was done to evaluate the system's persistence, recovery, and performance by extending existing tests and creating new ones.  After modifying the local RPC system to work on a real network, the system was deployed on an Amazon Web Services (AWS) cluster to obtain realistic measurements of throughput, latency, and disk recovery speed.

\section{Design} \label{sec:design}
The design is based upon the code developed throughout the semester for lab 4.  There are therefore three basic components running on each server: Paxos processes to reach consensus for instances, a shard master to determine which groups are responsible for various shards of the database, and a shard key/value storage process to execute client Put/Get requests.  The Paxos instances are used to create a consistent log of operations for both shard reconfiguration and client operations.

Battery tests were run to determine which combination of our prior labs created the fastest and most reliable starting point.  Once a working codebase was compiled, modifications were made to address three  major areas: Paxos optimizations, persistence (with disk recovery), and testing over a real network on AWS.

\subsection{Paxos Optimizations}
In order to improve the performance of Paxos, the system features the Multi-Paxos leader selection optimization. This optimization solved the dueling leaders problem, preventing infinite loops between conflicting leaders, and also reduced the number of RPC calls necessary to reach agreement by skipping the Prepare phase of the Paxos protocol for a leader that was constant between instances. The main issue with the Multi-Paxos leader selection is ensuring that leadership information is propagated through the system and ensuring that the protocol never agrees as part of two separate partitions when the Prepare phase is skipped. Unfortunately, a number of bugs were discovered with the pre-optimized Paxos implementation, which were fixed before producing a reliable Multi-Paxos optimized system that has a selected leader and skips unnecessary prepare phases.  Once completed, the number of required RPCs was reduced by about 33\%, and bandwidth requirements were reduced due to the singular leadership. There was an issue that a single machine became the primary proposer, potentially limiting the system based on its disk and processor speeds, however this doesn't appear to be an issue at the anticipated use cases.

\subsection{Persistent Storage}
In order to tolerate servers restarting and to handle databases that cannot fit in memory, all three components of the system store state and data persistently on disk.  To interface with the disk, a Go wrapper \cite{levigo} for the high-performance C++ key/value library LevelDB \cite{leveldb} was utilized.  An outline of what each system component stores on disk is presented below.
\begin{itemize}
\item The Paxos processes write the highest accepted value, the highest accepted proposal number, and the highest acknowledged prepare request to disk for each instance.  This ensures that the protocol will operate correctly even if a server restarts \cite{paxos} (for example, a decided value can never be changed).  In addition to this state, the Paxos processes persistently store the done responses received from peers and the number of the maximum known instance (see the recovery section below).

\item The shard master writes every configuration to disk, ensuring that will be able to respond to historical queries even after restarting.  In addition, it stores the highest executed Paxos log entry and the number of its highest known configuration.

\item The shard key/value service writes its database to disk by writing each value as each Put request is executed.  In addition, it must preserve at-most-once semantics and consistency across reboots, so it persistently stores which operations have been seen as well as the most recent response sent to each client.  It also writes the highest executed Paxos log entry and the current configuration number.
\end{itemize}

In all cases, the specified data is written to disk before returning to the caller, so any response that a caller receives from a peer will not be forgotten by the responder even if it restarts.  It may be the case that the responder crashes before sending the response, but this is acceptable since the system offers at-most-once semantics (so the caller must be prepared to retry and the responder is prepared to filter duplicates).

\subsubsection{Recovery from Failures}
In order to tolerate servers restarting with or without their disk contents, the system includes protocols for recovering and catching up on startup.  These can be summarized by considering what happens when a server starts:

\begin{itemize}
\item The Paxos process will load any saved done responses and maximum instance number from its disk.   It will then contact peers to request their done responses and maximum instance number, determining which peer is the most updated out of any peers that respond.  From these values, both the ones read from disk and the ones received, it can determine which instances it should know about but does not; it then requests these instances from its peers.
\item The shard master process will load the highest known configuration and highest executed Paxos log instance from disk if it exists.  It will then request the same information from its peers, determining who among the responders is the most updated.  From this information it can determine if there are any missed configurations, and if so request them from a peer.  
\item The shard key/value service will load its last known configuration and the highest executed Paxos log instance from disk if it exists.  It will then request the same information from peers, determining the most updated state that it should have.  Finally, it can request the actual database key/value entries from a peer (see below for details of how this transfer is performed) along with any client responses and lists of seen operations.
\end{itemize}
These protocols allow any number of servers restarting with their disk contents to catch themselves up from any updated peer.  They also allow any number of servers restarting without their disk contents to recover the needed data from any peer that has the data.

\textbf{Single server crashing and restarting, disk intact}: As described above, each
recovery process contacts a random replica in the same replica group.  The recovering server can then get any missing data by comparing states with the updated server (e.g. shard data, shard master configurations, and Paxos instances - note that Paxos only discards old decisions if all participants have marked that sequence number as 'done').

\textbf{Single server crashing and disk lost}:  This is essentially the same as if the disk was intact, except that much more data will need to be transferred (the same protocol can be used).  Note that when data is transferred, multiple messages may be used for a single shard if it is too large to fit in memory - the sender will periodically check memory usage while constructing the reply and mark replies as incomplete if needed.

% Is the below actually implemented?
%The recovery process contacts a random replica in the same replica group, and tells the up-to-date server that it has no state. The two servers open an auxiliary TCP connection to send over the relevant portions of the database. This avoids RPC overhead, since the database might be large. The database has an in-order iterator, in case the TCP connection fails, the recoverer can re-establish the connection and ask the sender to pick up where it left off. If the sending server fails, then the recovering server asks a new replica. This allows for a failure mode where the original server remains locked permanently. We have decided that this is preferable to sacrificing consistency, and the system administrator can resolve the problem by rebooting the hung server.

\textbf{Complete crash of all servers}: If all servers lose their disk state, then all will restart and begin like a brand new deployment (the above protocols will simply not find any updated data to transfer). If the disks are not lost, then the recovery process will read the various
states written to disk and pick up where it left off.  Because disk writes take place before any operation commits, data is not lost.

It should be noted that during a shard transfer (including recovery), the shard key/value service on both the sender and recoverer ignore client requests in order to ensure that consistent data is sent.  Since their Paxos processes continue running independently, however, progress can still be made even if a majority of servers is involved in a data transfer (the decided operations will be persistently logged and the halted servers can catch up when transfer is complete).

%% \subsubsection{Memory and Performance Considerations}
%% There are a number of design choices which were made in order to address both memory management and overall performance.

%% Firstly, the service must be able to handle database content which is too large to fit in volatile memory.  Towards this end, values are written to disk immediately as described above so it does not need to be kept in memory.  More importantly, however, servers that need to transfer shards (either because the shard master decided a shard needs to move or because a server is recovering) must be able to do so in multiple RPCs since memory is too small to create a reply with the entire database.  Therefore, a shard transfer occurs with the following protocol:
%% \begin{itemize}
%% \item The requester sends a request to a peer, including a request number, and the current configuration number.  If doing a shard transfer, the shard number is also included.
%% \item If the receiver is currently at the appropriate configuration, it accepts the request.  It will then begin populating a reply with client response data, a list of seen operations, and the desired key/value data.  While doing this, it will periodically check memory usage; if it exceeds a configurable threshold, it will reply with whatever it currently has and mark the reply as incomplete.
%% \item The requester copies the data from the reply to its disk.  If the reply was marked as incomplete, it sends another request to the same peer with an incremented request number.
%% \item This continues until the reply is marked as complete, at which point the requester launches a goroutine to make sure the sender receives an acknowledgement of the receipt of the final data message.
%% \item If at any point the requester decides that the contacted peer is dead, it launches an acknowledgement goroutine as previously described for that peer (to terminate the transfer in case the peer wakes up or is not dead) and moves on to try a different peer, resetting the request number to 0.
%% \end{itemize}
%% Both the requester and the sender ignore client requests until the transfer is complete - this ensures that the requester will receive a consistent snapshot of the data (and is the reason the requester launches acknowledgement goroutines since otherwise a peer erroneously declared dead would be halted forever).

%% One aspect which addresses performance is that progress can still be made while recovery is taking place as long as a majority of servers are not recovering.  For example, if a replica group has three servers and a shard transfer is occurring between two of them (perhaps because one just restarted), a client can still have its requests processed if it contacts the third peer.  This is because the Paxos service runs independently of the shard key/value service; although the latter will ignore client requests on the transferring servers, their Paxos processes can still respond and make progress.  Since the Paxos instances are stored persistently on disk, the two occupied servers can catch themselves up from the log when they complete the transfer.

%% In addition, effort was made to utilize memory in addition to the disk in order to increase performance.  The LevelDB database has the option to use a least-recently-used (LRU) cache to store recently used entries in memory.  In addition, a configuration variable was added to the service to allow values to be written to memory as well as disk until the memory is full.  These two options, combined with the native cache of the operating system, allow many needed values to be read quickly from memory rather than always consulting the disk.

%% When one server contacts another via the FetchRecovery RPC, both
%% servers enter a recovering state. This state means that neither are
%% allowed to modify their databases \emph{except} through the recovery
%% process. Depending on how many replicas there are, this may render a
%% replica group unable to process requests while recovery is underway.
%% If there are enough replicas so that paxos majority can still be
%% reached, then the recovering servers will catch up once they come back
%% online, using the functionality already built in to sync lagging paxos
%% servers.

%% All recovery processes send over both the relevant portions of the k/v
%% store, as well as paxos state and any state needed to detect
%% duplicates.

%% We use leveldb. To guarantee persistence, we enforce writing to disk
%% all operations before returning. This slows performance down, but
%% prevents loss of data that the client wants committed. Speed could be
%% gained using a Harp-like set of pointers, but this would be
%% complicated and the client would have to understand the modified
%% persistence model.

\subsection{Testing}
About 25 new tests were created, and existing ones were extended, in order to evaluate persistent storage and run on TCP sockets instead of unix file sockets.  A few flags and configuration variables can select different modes (e.g. with and without persistence, RPC using unix file sockets or localhost TCP).

Below are summaries of different located at 'mexos/src/\emph{$<$name$>$}/test\_test.go' unless otherwise noted.

\emph{paxos:} Paxos specific tests which include restarting servers deterministically or randomly.  These include cases where servers lose their disks, where all servers restart, where partitioning changes, where communication is unreliable, and many combinations of the above.  There are also benchmarks for measuring performance.

\emph{shardmaster:} Shardmaster specific tests which include restarting servers with or without disk to ensure that persistence and recovery work as expected.  There are also benchmarks for measuring performance.

\emph{shardkv:} Shardkv specific tests which include deterministically restarting servers to ensure that persistence and recovery operate correctly.  There is also a test which performs the previously written tests but with random reboots of random servers or entire groups.  There are also benchmarks for measuring shard transfer speed and to ensure that the simulated memory limit is never exceeded.

\emph{test:} Tests and benchmarks modified to run on our AWS cluster.  Running these tests are the most complicated, and require the user to SSH into the AWS cluster leader and start the servers on other cluster machines using 'start-on-all.sh' to run 'mexos/src/main/start.go'. These tests required a surprising amount of debugging and glue code.  Additionally, to check for correct operation under server failure, backdoor RPCs were required to tell a remote server to act deaf or dead.

The results of these benchmarks are presented below.

\section{Performance}

An extensive testing system was created to try and accurately measure performance.  This comprises many new benchmarks, as well as a modified RPC system to run the servers in an AWS cluster and observe the performance on a real network.

\subsection{Throughput and Latency}

Performance tests were run using both vanilla paxos and multipaxos under a variety of conditions. The tests were run on an actual network in an AWS cluster.

\textit{Client latency}: Average time it takes for one client to see a
request processed.

\textit{Throughput}: Number of requests, on average, the
client will see processed in a second.

\textit{System throughput}: Number of requests per second the entire system can handle. This is different than the client throughput because clients always wait to see the response of the current request before dispatching another.

System throughput requires a saturation of multiple clients sending continuous requests for a certain amount of time (here, ten seconds). The system saturated from 10 to 50 clients, depending on the number of groups/replicas. After the saturation point, adding more clients
causes a decrease in throughput because of network traffic saturation.

Some tests were repeatedly run on one key, and others on a random set of keys. This helps indicate how the performance changes when requests all go to the same replica group versus when they are evenly spread across all replica groups.

% Please add the following required packages to your document preamble:
% \usepackage[table,xcdraw]{xcolor}
% If you use beamer only pass ``xcolor=table'' option, i.e. \documentclass[xcolor=table]{beamer}
\begin{table}
\begin{tabular}{|l|l|l|l|}
\hline
\multicolumn{4}{|c|}{\cellcolor[HTML]{C0C0C0}{\color[HTML]{000000} \textbf{Vanilla Paxos}}}                           \\ \hline
                              & \textbf{Avg. Client Latency} & \textbf{Avg. Client Throughput} & \textbf{System Throughput} \\ \hline
\multicolumn{4}{|c|}{\textit{PutHash to one key}}                                                                     \\ \hline
\textbf{1 group, 3 replicas}  & 40.3 ms/op                   & 24.8 requests/s              & 27 requests/s           \\ \hline
\textbf{1 group, 10 replicas} & 5370 ms/op                   & 0.18 requests/s              & 2 requests/s            \\ \hline
\textbf{3 groups, 3 replicas} & 62 ms/op                     &                              & 22 requests/s           \\ \hline
\multicolumn{4}{|c|}{\textit{PutHash to many keys}}                                                                   \\ \hline
\textbf{1 group, 3 replicas}  & 40.7 ms/op                   & 24.5 requests/s              & 30 requests/s           \\ \hline
\textbf{1 group, 10 replicas} & 282 ms/op                    & 3.5 requests/s               & 7 requests/s            \\ \hline
\textbf{3 groups, 3 replicas} & 214 ms/op                    & 4.6 requests/s               & 67 requests/s           \\ \hline
\end{tabular}
\end{table}

% Please add the following required packages to your document preamble:
% \usepackage[table,xcdraw]{xcolor}
% If you use beamer only pass ``xcolor=table'' option, i.e. \documentclass[xcolor=table]{beamer}
\begin{table}
\begin{tabular}{|l|l|l|l|}
\hline
\multicolumn{4}{|c|}{\cellcolor[HTML]{C0C0C0}{\color[HTML]{000000} \textbf{Multipaxos}}}                              \\ \hline
                              & \textbf{Avg. Client Latency} & \textbf{Avg. Client Throughput} & \textbf{System Throughput} \\ \hline
\multicolumn{4}{|c|}{\textit{PutHash to one key}}                                                                     \\ \hline
\textbf{1 group, 3 replicas}  & 27 ms/op                     & 37 requests/s                & 47 requests/s           \\ \hline
\textbf{1 group, 10 replicas} & 2360 ms/op                   & 0.42 requests/s              & 3.4 requests/s          \\ \hline
\textbf{3 groups, 3 replicas} & 22 ms/op                     & 45 requests/s                & 45 requests/s           \\ \hline
\multicolumn{4}{|c|}{\textit{PutHash to many keys}}                                                                   \\ \hline
\textbf{1 group, 3 replicas}  & 25 ms/op                     & 40 requests/s                & 45 requests/s           \\ \hline
\textbf{1 group, 10 replicas} & 130 ms/op                    & 7 requests/s                 & 10.3 requests/s         \\ \hline
\textbf{3 groups, 3 replicas} & 40 ms/op                     & 25 requests/s                & 124 requests/s          \\ \hline
\end{tabular}
\end{table}

It can be seen that, in all cases, the Paxos optimizations had a significant positive impact on both latency and client and system throughput.  For example, system throughput increased by an average of over 71\%.

For a single group case, the difference between three replicas and ten is striking.  With more replicas, more traffic is needed for a paxos agreement and thus latency increases.  Figure 1 demonstrates how throughput suffers as the number of replicas increases.  Another interesting effect is that as the number of replicas increases, the gap between the single key and multi-key cases increases dramatically even when there is a single group (and therefore all puts touch the same servers).

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{
\includegraphics[scale=.40]{replicas.png}}
\caption{Performance of a one-group configuration rapidly deteriorates as the number of replicas increases}
\label{fig:replicas}
\end{figure}

Adding more groups, however, increases throughput.  For evenly distributed traffic, going from one group to 3 nearly triples the throughput when using multiple keys (thereby leveraging the additional groups).  Figure 2 shows that this trend continues, such that adding groups linearly improves the throughput of the system. The table shows that average throughput for a single client may suffer, probably because the system throughput benchmark creates many clients, while the client throughput test creates a single client; having multiple clients increases the likelihood that a client must wait to see its request served since the replicas may be busy serving another client's request.

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{
\includegraphics[scale=.4]{groups.png}}
\caption{Performance increases linearly as the number of groups increases. There is probably a saturation point, but we didn't have enough cluster machines to reach it.}
\label{fig:groups}
\end{figure}

\subsubsection{Paxos RPC Counts}

Utilizing optimizations similar to Multipaxos decreases the number of RPCs needed to reach agreement.

% Please add the following required packages to your document preamble:
% \usepackage[table,xcdraw]{xcolor}
% If you use beamer only pass ``xcolor=table'' option, i.e. \documentclass[xcolor=table]{beamer}
\begin{table}[h]
\centering
\makebox[\textwidth][c]{
\begin{tabular}{|l|l|l|}
\hline
\multicolumn{3}{|c|}{\cellcolor[HTML]{C0C0C0}\textbf{Paxos RPC Counts}}                  \\ \hline
                                 & \textbf{Single Proposer} & \textbf{Multiple Proposers} \\ \hline
\textbf{Normal Paxos}            & 30                       & 89            \\ \hline
\textbf{Paxos with Leaders}      & 20                       & 27                         \\ \hline
\textbf{Leaders and Pre-prepare} & 17                       & 20                         \\ \hline
\end{tabular}
}
\end{table}

\subsubsection{Disk Recovery}
Various tests were performed to evaluate the recovery time for servers restarting.  The AWS servers can either have hard disk drives or solid state drives, and both were evaluated for raw disk speed by writing 1GB of data to the database.  A HDD averaged about 47.7 MB/s while the SSD averaged about 166.6 MB/s.  The recovery was then tested by creating a one-group service, placing 100 MB worth of data in the database, and finally restarting a replica without its disk contents.  Various tests were run with differently sized values (and therefore differing numbers of entries).  In all tests, the memory limit was set to under 75 MB, so that data transfer was forced to take place using multiple messages.

\begin{table}
\begin{tabular}{|l|l|l|l|}
\hline
\multicolumn{4}{|c|}{\cellcolor[HTML]{C0C0C0}{\color[HTML]{000000} \textbf{Recovery Performance for 100MB Database}}}                           
\\ \hline \textbf{Value Size (KB)} &  \textbf{Number of Entries} & \textbf{Time Reading Database (s)} & \textbf{Recovery Time (s)} 
\\ \hline 5  & 20352 & 464.7 & ?
\\ \hline 50  & 2046 & 30.8 & 37.2
\\ \hline 100  & 1023 & 15.9 & 21.1
\\ \hline 150  & 682 & 10.2 & 15.8
\\ \hline 250  & 409 & 6.3 & 12.4
\\ \hline 500  & 204 & 3.5 & 10.6
\\ \hline 1024  & 99 & 1.88 & 8.3
\\ \hline 1500  & 68 & 1.4 & ?
\\ \hline
\end{tabular}
\end{table}

\begin{figure}[h]
\centering
\makebox[\textwidth][c]{
\includegraphics[scale=.4]{recovery.png}}
\caption{Performance increases dramatically as the number of keys is reduced (and the size of entries is increased).}
\label{fig:recovery}
\end{figure}

In these experiments, an average of about 5 messages was sent in order to complete the shard transfers.  In the above table, the database read time is the total time that the sender spent scanning its database to collect the data needed to be sent.  The remaining time is spent transferring the data over the network and by the receiver writing a copy to disk.  It can be seen that this time accounts for the vast majority of total transfer time when there are many keys, and decreases dramatically as the number of entries is reduced.  This indicates that the main bottleneck is iteration over the database at the sender.  The current implementation uses the built-in LevelDB iterator - future research can be done into how this operates (and into whether the database is fragmented on the drive) to try and improve performance since the aforementioned measured disk speeds indicate that it should be able to read the data much faster.  Meanwhile, the networking time remains approximately constant regardless of the number of entries, indicating that the network bandwidth itself is the limiting factor for RPC time.  The fact that the number of messages remained approximately constant indicates that the recovery protocol does not have significant overhead which depends on the number of keys but rather is simply limited by memory capacity as desired.  As the number of keys decreases, the network therefore becomes the limiting factor.

%Raw time to write 1GB to database, HDD:  21.470750409s

%Raw time to write 1GB to database, SSD:  6.1470524260s

%Recovery time of a 1GB database using our system: TBD

\subsection{Bottlenecks}
Although the above enhancements significantly improved performance, various were identified which could be addressed in the future.

One notable bottleneck is disk reading and writing.  The database library wrapper, LeveDB, offers a cache as well as compression and is quite fast.  However, as discussed previously, benchmark tests revealed that scanning the database to compile data for shard transfers accounted for a significant portion of total transfer time.  Additionally, in the process of debugging networked RPCs, simple logging was implemented which slowed down operations by nearly a factor of ten; this was a good demonstration of disk latency and how care must be taken in even mundane aspects of the system.  It was considered adding a commit point marker (similar to the ideas presented in \cite{harp}), which would be straightforward in the current system by simply returning before the disk write and not advancing the Paxos min until the write completes, but the above measurements indicate that database scanning is much more significant than isolated accesses.  In the future, the improvements can be made to avoid this iteration bottleneck.  For example, the system topology could be altered to allow for more parallelizable recovery with a procedure inspired by FDS \cite{fds}.  More simply, however, the servers can be altered to store a separate database for each shard; the entire database could then be transferred to the receiver using TCP/UDP.  Preliminary tests on AWS showed that a 100MB database can be sent over TCP in less than one second.  In addition, this would avoid the need for either the sender or the receiver to scan through a database, thus eliminating the major bottleneck identified previously.

Another bottleneck is network capacity and/or number of messages.  As shown in the latency and throughput section, increasing the number of replicas has an adverse effect on performance since each operation requires communication with a majority of replicas.  The paxos optimizations therefore helped to improve performance, and further optimizations could be made to the data transfer protocols to reduce the required number of messages.

Increasing the number of replica groups increases the performance of the system for normal operation, especially under heavy load with evenly distributed key requests. This is especially noticeable in the system throughput metric.  This arises since number of groups roughly corresponds to how parallelizable the operations are.

Using multipaxos reduces the amount of network communication required, which improves the throughput and latency by about a factor of two. Multipaxos also saves some disk space by reducing the amount of stored paxos state information since prepares are pre-determined.

It is also suspected that using UDP would improve throughput, as servers that do not receive a response already re-transmit their request, making the TCP acknowledgement system redundant.

\section{Conclusion}
This project implements, tests, and benchmarks a persistent and performant sharded key/value store.  While there are various improvements and further testing that were not yet implemented, the current system comprises a working model which demonstrates the effectiveness of its various protocols and optimizations while revealing the bottlenecks that can be addressed in the future.

%{\footnotesize \bibliographystyle{acm}
%\bibliography{paper}}
%\newpage
\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,paper}
\end{document}